Palomar 5 Gaps

### 2024-10-02 paper/main.tex
passed over results with GPT

    Firstly, to our knowledge this is the first time any one tried to study the interactions between the globular clusters. By including N~body forces on the orbits with the globular clusters, we did not observe any significant changes to orbital properties such as energy or angular momentum between any of the globular clusters in any of the simulations. This is an expected result given their small sizes and vast distances between them.

    We inspected all simulations upon their completion and present one of them in in Fig.~\ref{fig:stream_on_sky}--which we selected for its clear and prominent gaps. Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitutde to view it's 1D profile density over longitude.

    We apply the gap detection methods described in Appendix.~\ref{sec:gap_detection} and the perturber identification methods from Appendix.~\ref{sec:Perturber_Identification}. We present the results from all the simulations in Appendix.~\ref{sec:gallery_of_gaps}. Since we identify which perturber is responsible for a given gap, we launch the more simulations in which only one globular cluster at a time is included to isolate specific gaps. We present this in Fig.~\ref{fig:decomposition}.

@@

    \begin{figure*}
      \centering
      \includegraphics[width=\linewidth]{histogram_impact_time.png}
      \caption{The time of impact for all gap causing flybys summed over all 50 simulations. Each perturbing cluster is label and is color consistent. Time is presented in integration units. Note that 1~s~kpc~km~$^{-1}$ is roughly 1~Gyr. Also, the y-axis is broken to show the total number of encounters from NGC~2808 yet not dwarf the scale from the other interactions.}
      \label{fig:histogram_impact_time}
      \end{figure*}

    A striking result from the simulations is that more than half of the gaps are caused by NGC~2808. Upon inspecting the figures in the appendix, we notice that these gaps are of similar size and occur in nearly the same position in the leading tail across multiple simulations. The size of a gap is a function of time, and recent encounters leave limited time for the gap to grow. As shown in Fig.~\ref{fig:histogram_impact_time}, the gaps from NGC~2808 do occur at about 200 Myr ago. It is no coincidence that a recent fly occurs in most of the simulations. As we look further back in time, present-day uncertainties allow for much broader variation in the phase space positions of the globular clusters. However, the uncertainties associated with NGC 2808 and Palomar 5 are sufficiently low that the recent flyby of NGC 2808 occurs in 44 out of 50 simulated cases. The flybys do not happen exactly the same way in each simulation, the range in contrast of the gaps shown in the appendix illustrates that some encounters were closer, leading to more significant impacts.

    Next, we notice that most of the gaps occur on the leading tail. However, we believe this to be a coincidence. Stars escape from the globular cluster at the lagrange points at roughly equivalent rates. The two substantial differences are the orbital phase, and that the leading tail is about ~400~pc closers to the galactic center than the trailing tail. However, this is peanuts and negligible compared to the orbital radius of 8-12~kpc. Therefore, we expect the probability of a gap appearing in the leading or trailing tail to be equal and the presence of one gap should not influence the likelihood of another.

@@

    Overall, in 50 simulations we observe 73 gaps. The distribution of number of simulations with a give number of gaps is zero: 4, one: 25, two: 16, three: 4, four:1. If we regard NGC2808 as a fluke, then we are stil left with 29 gaps. Thus, in our simulations we expect [0.6-1.5] gaps.

>>>>

    To our knowledge, this is the first study to investigate interactions between globular clusters in this context. We included N-body forces on the orbits with the globular clusters, yet did not observe significant changes to orbital properties, such as energy or angular momentum, between any of the globular clusters in any simulation. This result is expected given the small sizes and vast distances between the clusters.

    We inspected all simulations upon their completion and present one particularly clear example shown in Fig.~\ref{fig:stream_on_sky} which we selected for its prominent gaps. Two of these gaps are visible in galactic coordinates and become even more apparent when marginalizing over latitude to view their 1D density profile over longitude.

    We applied the gap detection methods described in Appendix~\ref{sec:gap_detection}, along with the perturber identification methods from Appendix~\ref{sec:Perturber_Identification}. Results from all simulations are presented in Appendix~\ref{sec:gallery_of_gaps}. Once we identified the perturbers responsible for each gap, we conducted additional simulations to verify correct identification of the perturbers. These additional contain only the suspected clusters and were launched one-at-a-time to isolate the effects of a specific clusters. These results are shown in Fig.~\ref{fig:decomposition}.

@@

    \begin{figure*}
      \centering
      \includegraphics[width=\linewidth]{histogram_impact_time.png}
      \caption{The time of impact for all gap causing flybys summed over all 50 simulations. Each perturbing cluster is label and is color consistent. The time axis is given in simulation units, with 1~s~kpc~km~$^{-1}$ corresponding to roughly 1~Gyr. Note that the y-axis is broken to accommodate the large number of encounters from NGC2808, without overshadowing the other interactions.}
      \label{fig:histogram_impact_time}
      \end{figure*}

    One striking result is that more than half of the gaps observed are caused by NGC2808. Upon examining the figures in the appendix, we note that these gaps are of similar size and occur in almost the same position in the leading tail across most simulations. This is no coincidence, as the size of a gap is a function of time, and recent encounters allow only limited time for the gap to develop. As shown in Fig.\ref{fig:histogram_impact_time}, the gaps caused by NGC2808 result from a flyby that occurred approximately 200 Myr ago. It is expected that a recent flyby appears in most simulations, as uncertainties grow when we look further back in time, allowing broader variation in the phase space positions of the clusters. However, due to relatively low uncertainties associated with NGC2808 and Palomar 5 compared to the integration time, the recent flyby of NGC~2808 occurs in 44 out of 50 simulations. Although the flybys do not occur identically in each simulation, the range in gap contrast, shown in the appendix, indicates that closer encounters lead to more significant impacts.

    Additionally, we observe that most gaps occur in the leading tail. We believe this to be a coincidence, as stars escape from the globular cluster's Lagrange points at roughly equal rates. The two main differences are the orbital phase and that the leading tail is about 400~pc closer to the Galactic center than the trailing tail. However, this difference is negligible compared to the orbital radius of 8-12kpc. As such, we expect the likelihood of a gap appearing in either tail to be equal, with the presence of one gap not influencing the probability of another.

@@

    Overall, across 50 simulations, we observe 73 gaps. The distribution of simulations with a given number of gaps is as follows: 0 gaps in 4 simulations, 1 gap in 25 simulations, 2 gaps in 16 simulations, 3 gaps in 4 simulations, and 4 gaps in 1 simulation. Excluding NGC~2808, we are left with 29 gaps, leading to an expected range of 0.6–1.5 gaps per simulation.
